Using WordNet and Short-Term Memory for Contextual Disambiguation

### 2017-04-04 Draft_2.tex
Finished first draft, and began work on second draft

        \Huge
        \textbf{Using WordNet and Short-Term Memory for Contextual Disambiguation}
        \vspace{2cm}

@@

        \Large
        Word Count - 4507\\
        Supervisor - Dr Dimitar Kazakov\\
        Department of Computer Science\\
        University of York\\
        20 November 2016

    \end{center}
\end{titlepage}

@@

\section{Introduction}
\label{sec:Intro}

\section{Literature Review}
\label{sec:LitReview}
In order to define a problem, we must first establish the previous works, so as to build upon them effectively.

%--------------------- NLP

\subsection{Natural Language Processing}
\label{sec:NLP}
The study of natural language processing aims to allow a computer to understand natural language, and formulate a relevant response based upon its input. Within this, the problem of input text analysis has traditionally be broken down into smaller sub-problems\cite{NLPHandbook}:
\begin{itemize}
	\item Text Preprocessing
	\item Lexical Analysis
	\item Syntactic Parsing
	\item Semantic Analysis
\end{itemize}
The subsequent subsections will discuss each of these in more detail.

\subsubsection{Text Preprocessing}
\label{sec:TextPreprocessing}
Before any analysis can take place, the inputted raw text must be converted into a usable format. This, once again, can be broken down into multiple steps\cite{NLPHandbook}:
\begin{itemize}
	\item \textbf{Document Triage}

	\begin{itemize}
		\item Character encoding must be identified.
		\item The language can then be identified. %Language is most commonly found using one of two methods: Language can be identified by character set, in cases where language uses a unique alphabet, or by character frequency, in other cases.
		\item Non-useful data, such as images and html formatting must be removed.			\end{itemize}	
	\item \textbf{Text Segmentation}

	\begin{itemize}
		\item Individual words (tokens) must be separated from one another. % This is done using white space in space-delimited languages, and comprehensive lists in unsegemented languages
		\item Text Normalisation; replacing multiple equivalent tokens with one token (e.g. "Ave." and "Avenue").
		\item Identifying sentences, i.e. locating where a sentence begins and ends. 
	\end{itemize}

\end{itemize} 
% ########################
% #  UNFINISHED SECTION  #
% ########################

\subsubsection{Lexical Analysis}
\label{sec:LexicalAnalysis}
One word can have multiple forms, for example "judge" (the lemma) has the forms \{"judge", "judges", "judging", "judged"\} (morphological variants).  The job of Lexical Analysis is to replace all morphological variants of a word, with their corresponding lemma, a process known as stemming \cite{NLPHandbook}.

% In terms of ease of token identification, it can be seen conceptually how this process is beneficial. With that said, when we consider the amount of information in a word, it can similarly be realised that we lose information about the form of a word, for example plural/singular and tense.
\subsubsection{Syntactic Parsing}
\label{sec:SyntacticParsing}
Whwn derviving meaning from sentances, the grammatical structure can provide important insight. The Syntactic parsing technique extracts this infomation using two processes \cite{NLPAlmostFromScratch}: 
\begin{itemize}
	\item \textbf{Part-of-speech (PoS) Tagging}
	\label{PoSTag}
	\begin{itemize}
		\item Each word is given tags denoting their syntactic role (e.g. noun/adjective/verb).
	\end{itemize}

	\item \textbf{Chunking}

	\begin{itemize}
		\item Noun phrases and verb phrases are detected and tagged using "begin chunk" and "inside chunk" labels
	\end{itemize}

\subsubsection{Semantic Analysis}
\label{sec:SemanticAnalysis}
The semantics of a sentence, is the meaning given by it's tokens \cite{SemanticAnalysisAPracticalIntro}. The topic of semantic analysis will be expanded upon on in the \hyperref[sec:PrevWork]{Previous Works Section}.

@@

\subsection{Project Aims}
\label{EvAims}
In section \ref{sec:ProjectAim}, three aims for the project were established.

\subsubsection{A computational model of memory structures}
\label{EvAims1}
The memory structures implemented are modelled on those described by Baddeley \cite{MemoryBaddeleyEysenkAnderson}.

\subsubsection{A system for reading in a corpus}
\label{EvAims2}

\subsubsection{A computational model of disambiguation}
\label{EvAims2}

\section{Conclusions}
\label{sec:Conclusions}

>>>>

        \Huge
        \textbf{Using WordNet and a Short-Term Memory model for Contextual Disambiguation}
        \vspace{2cm}

@@

        \Large
        Wordcount - 9191\\
        (Wordcount produced using wc - w)\\
        Supervisor - Dr Dimitar Kazakov\\
        Department of Computer Science\\
        University of York\\
        20 November 2016

    \end{center}
\end{titlepage}
\
\vfill
\begin{center}
	\textbf{Abstract}
\end{center} 
\label{sec:Abstract}
Semantic Analysis is the process of giving meaning to the tokens of a sentence. This process has a multitude of applications from speech recognition to unsepervised knowledge aquisition. Current solutions do not provide enough accuracy to be relied upon fully for these puposes, making it a heated area o research. It is known that humans are very good at semantic analysis, and multiple psychologial models describing this exist, such as the Working Memory Model. The focus of this report is a computational model, derived from these theories, with the intention of providing a more reliable solution to the problem of semantic analysis. A subset of the working memory model is used, in conjunction with hyponymy base activation and disambiguation. The model developed provides an acccuracy greater than some approaches, though, without further development, fails to reliably identify word meanings across a large input corpus. The results implicate the need for the use of a greater number of semantic links in order to correctly derive meaning.

\newpage

@@

\section{Introduction}
\label{sec:Intro}
%--------------------- NLP

\subsection{Context}
\label{sec:IntroContext}

Natural language is the language spoken and written by humans. It is very powerful, though its key downfall is its ambiguity. The study of natural language processing aims to allow a computer to understand natural language, and formulate a relevant response based upon its input. Within this, the problem of input text analysis has traditionally been broken down into smaller sub-problems \cite{NLPHandbook}:

\begin{itemize}
	\item \textbf{Text Preprocessing} \\
	Before any analysis can take place, the input raw text must be converted into a usable format \cite{NLPHandbook}.
	\item \textbf{Lexical Analysis} \\
	One word can have multiple forms, for example ”judge” (the lemma) has the
	forms {”judge”, ”judges”, ”judging”, ”judged”} (morphological variants). The job of Lexical Analysis is to replace all morphological variants of a word, with their corresponding lemma, a process known as stemming \cite{NLPHandbook}.
	\item \textbf{Syntactic Parsing} \\
	When deriving meaning from sentences, the grammatical structure can provide important insight. The Syntactic parsing technique extracts this information using Part-of-Speech tagging (detecting syntactic role) and Chunking (detecting noun and verb phrases) \cite{NLPHandbook}.
	\item \textbf{Semantic Analysis} \\
	The derivation of meaning from the sentences tokens \cite{NLPHandbook}. 
\end{itemize}

Semantic Analysis is a problem which continues to draw attention from researchers, due to the lack of a truly capable solution \cite{NLPHandbook}. Most would agree that, though a good computational solution is not available, the human brain performs impressively well when confronted with the same problem. Given the example:

\[The\: party\: led\: government\]

it is known by the reader that “party” refers to a political party, due to the fact that it is leading government. The meaning of the word is taken from its surrounding context.

It can be seen that the study of the brains processes of semantic analysis, and their application to a computational model, could provide a satisfying solution. A great deal of research is directed at studying these process, largely influenced by the working memory model proposed by Baddeley \cite{MemoryBaddeleyEysenkAnderson}. This memory model describes a system of structures, which work in conjunction with one another to form a representation of the brain’s input.

The Working Memory model has many more applications than the one described in the coming report, requiring more structures than are relevant to the process of semantic analysis. For the purposes of this application, we will restrict our focus to three of these memories, short-term, long-term, and the Episodic buffer\cite{MemoryBaddeleyEysenkAnderson}; each of which is described in greater detail in the Literature review.

The following report will discuss and define a potential computational model of the psychological theories surrounding the process of Semantic Analysis. As part of the project, a system will be built, and tested against an input corpus. Due to time, and performance constraints, a level of abstraction will be required. I is the purposes of this testing to decide whether the level of abstraction used is provides a satisfactory performance.

\subsection{Ethics}
\label{sec:IntroEthics}

The project conducted as part of this report requires no outside participants, nor does it directly contribute to any research or action that could lead to harm. The purposes of this project are purely research based, with the intention of leading to further research in the same field. It is assumed that previous research used by this project was conducted following generally accepted ethical practices.

\section{Literature Review}
\label{sec:LitReview}

@@

\subsection{Overall Evaluations}
\label{EvOverall}
The system as a whole has offered an improvement over frequency analysis, with an accuracy of 40\% compared to 12\%. All but one aspect of the system provided some benefit to the process of semantic analysis, with \hyperref[sec:EvSanity]{Saity Checking} unfortunately failing to offer any improvement.

It can be seen, when comparing testing data (one document analysed) to evaluation data (five documents analysed), that the system performed significantly better over some documents than others. This difference is likely due to the level of ambiguity in the input text, with a news piece being used for testing, and topics covering opinion and fiction being used for evaluation.

\section{Conclusions}
\label{sec:Conclusions}
This project has aimed to utlise previously established psychological models of disambiguation, and apply them to the problem of Semantic Analysis. The models used are those originally developed by Atkinson and Shiffrin \cite{ControlProcessesSTMAtkinson}, and then later further studied by Baddeley \cite{MemoryBaddeleyEysenkAnderson}.

\subsection{Project Aims}
\label{ConcAims}
In section \ref{sec:ProjectAim}, three aims for the project were established. All three aims have been met, with varying degrees of success.

\begin{enumerate}
	\item \textbf{A computational model of memory structures} \newline
		As part of this project, three memory structures have been produced, STS, LTS and the Episodic Buffer.
		\begin{itemize}
			\item The STS is a memory structure capable of holding a finite amount of data regarding the local context of a sentence. In the implementation of this project two STSs were used, one for verbs and the other for nouns. This decision does not follow the Badeley's model, in which there exists only one STS, containing data regarding all word types \cite{MemoryBaddeleyEysenkAnderson}. With that said, words are able to move to and from the STS, each with its own activation level.

			\item The LTS, provided by WordNet, contains a large number of synsets connected to one another by semantic links. Hyponomy was the most heavily used of these links, though others such as meronymy, were not used. The functions of nouns were represented using separate dictionaries, due to the fact that WordNet does not yet contain this information, though these dictionaries were far from complete.

			\item The Episodic Buffer was shown in testing to have a significant effect on accuracy. This is likely due to its ability to correctly represent the global context of a document. In this implementation, items would never leave the episodic buffer, meaning that synsets that appear early in the document, can still affect activation, long after context has changed.
		\end{itemize}

	\item \textbf{A system for reading in a corpus} \newline
		The system uses atctivation and forgetting to model the same prrocesses which exist in Atkinson and Shiffrin's model \cite{ControlProcessesSTMAtkinson}. As the input corpus is read, these processes occur for each word, causing context to be represented by the STS and Episodic Buffer.

	\item \textbf{A computational model of disambiguation} \newline
		The disambiguation model makes heavy use of the contents of the STS, in conjunction with the hyponomy relation. Frequency is also used in cases where the correct synset can not be identified using context alone, fitting with the Multiple Access Model \cite{PsychologyOfLanguage}. Unfortunatly, the disambiguation system only produced an accuracy of 40\%, which, though significantly better than frequency analysis alone (12\%), is not good enough to be directly useful.

The seemingly low accuracy of the system does not imply that it is is of not of use. We can see that it is capable of finding the correct synset in some cases. In others, it is not unlikely that the system reduced the number of possibilities to a smaller subset, and from then, failed to select the correct option. In this case, the described implementation could be used to reduce the number of possibilities, which in turn could be made smaller still by some future extension.

\subsection{Future Extensions}
\label{sec:ConcFuture}

\subsubsection{Semantic Links}
\label{sec:FutureSemantic}
In this implementation, some semantic links have not been utilised. One of such links is Meronymy. In the \hyperref[sec:ProbDef]{Problem Definition section}, the following example was given:

It was stated that the brain can identify that "party" refers to a political party, due to its use with the word government. This is an example of where meronymy is used, with a government containing multiple parties. "Party" and "government" share no common hypernyms in WordNet (excluding "Synset{entity}"), meaning the system, as described in section \ref{sec:Implementation}, would fail on this sentence.

Another semantic relation partially excluded from this implementation is functions. An attempt at using this semantic link was made, with the inclusion of \hyperref[sec:Sanity]{Sanity Checks}. Unfortunately, The Sanity checking system was detrimental to the system as a whole, most likely due to the limited amount of data available to it.

Functions are a proposed feature of WordNet \cite{WN2Nouns}, though due to the scale of the inclusion, they are yet to be added to the available database. Inclusion of noun functions would reduce the need for independent memory structures for nouns and verbs. This relation would allow locally used nouns to contribute to the disambiguation of verbs, and vice versa, an especially useful process in cases where noun/verb based context is low.

\subsubsection{Improved Usage of Syntactic Parsing}
\label{sec:FutureSyntactic}
Syntactic parsing, more specifically chunking, provides information about verb and noun phrases. These phrases provide a more specific local context than a set of sentences, as used by the implemented system.

Noun phrases contain a noun and its related adjectives. In this case, the adjectives used can provide some indication of the meaning of the noun, for example:

\[The\: white\: cloud.\]

In this noun phrase, we know that "cloud" does not refer to the internet, because the internet has no colour. This extends to verb phrases, where the set of possible meanings could be reduced using functions, as discussed in the previous subsection.By using knowledge of which words directly relate to one another in a sentence, we know which context is relevant, and which is not.
